---
title: "lab5_bf"
author: "Bret Folger"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)
```

## Objectives

 - Getting counts for different grouos
 - Use (lubridate) to parse dates
 - Find confidence intervals & do t-tests witgh t.test()
 - Make our first heatmap with geom_tile()
 
 
```{r, include=FALSE}

#Import lobster data, get rid of NAs, clean names


lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"),
                              na = "-99999") %>% 
  clean_names()
  

```

Use the tidyr::uncount() function to expand a frequency variable

```{r}
lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)

```


Yay! Now each lobster has its own row. This is in tidy format

Let's look at this a bunch of ways:

```{r}
ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) + 
  geom_jitter(aes(color = site),
              width = 0.2,
              alpha = 0.5) 

ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) + 
  geom_violin(aes(color = site)) 

ggplot(data = lobster_tidy, aes(sample = size_mm)) + 
  geom_qq() +
  facet_wrap(~site, scales = "free")
  
```

They all look pretty normally distributed!

Use (lubridate) package to parse dates and times
When using lubridate, specify that you are using the lubridate package because the functions within it can be found elsewhere

Add a new column using mutat that contains my date as an actual date:

```{r}

#important to know the beginning format of the date. here it is mdy

#call the date column at the end of "mdy()" in order to get it int he right format

lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )



```

Now we will parse 'date' column to get the different pieces (month, year) separated

```{r}

#label = TRUE gets the month abbreviation in there

lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE)
  ) %>%
  mutate(
    obs_day = lubridate::day(date_new) 
  ) %>% 
  mutate(
    obs_year = lubridate::year(date_new)
  ) 
  

```

Count lobsters by different groupings...

Using "count()" from dplyr. Groups for you then ungroups when you're done. Amazing

Let's count the number of lobsters by year and month

```{r}

lobster_ym <- lobster_parse_date %>% 
  count(obs_year, obs_month)

#now lets count by lobster year and then site

lobster_ysite <- lobster_parse_date %>% 
  count(obs_year, site)

lobster_ysite

lobster_site <- lobster_parse_date %>% 
  count(site)

lobster_site

lobster_length <- lobster_parse_date %>% 
  count(size_mm)

lobster_length

# group_by + summarize + n()
#n uses the groups you previously denoted to get the sample size

lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    sample_n = n()
    
  )

```

NEXT: Confidence intervals and t-tests

Use the 't-test()' function to find confidencef intervals and perform t-tests

```{r}

ivee_lobsters <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobsters)

```

Two - sample t-test: 

Is there a significant difference in mean lobster lengths between naples reef and mohawk reef?

```{r}

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mohk_napl_ttest <- t.test(napl_sample, mohk_sample)

mohk_napl_ttest

```

What does the p-value mean? Read about it - check out Khan Academy

Inline referencing...(whatever that is):

Mean lobster size differed significantly between Mohawk and Naples reefs. (t(`r mohk_napl_ttest$parameter`) = 19.849)

Maybe an easier way to do a 2 sample ttest:

```{r}
lobster_2sample <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

#this maxes out at 2 groups for the site column

ttest_2 <- t.test(size_mm ~ site, data = lobster_2sample)

ttest_2

```

Now let's make a geom_tile heat map

```{r}

ggplot(data = lobster_ysite, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))

```

